### 2021-12-01 plos_latex_template.tex
RJ: Edits to rule 4

% Rocío: Idea of this paragraph: non-transferability and do not restrict people
People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. The experience of people in marginalized groups facing barriers for equal participation and trying to solve them are invaluable; they cannot be replaced by good intentions from people that have not been through the same experiences \cite{costanzachockDesign2020}.
% People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.%ast maybe frame this on a more positive note (ideas, problem solving)
However, do not require nor expect minoritized people to only work—or want to work—on diversity and inclusion aspects; 
everybody must have the freedom to choose in which areas of the conference they want to work. 
% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
Furthermore, do not assume that self-nomination and voting will work as mechanisms to choose leadership positions. Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
Build an environment in which every person can express their opinion and pay special attention to people from systematically excluded groups.
Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget as well as other types of recognition (e.g. public recognition on social media, on the conference website, or during the conference).
In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.

@@

When choosing or inviting people for visible and valued roles in the conference––such as keynote speakers, program committee members, session chairs––, it is likely that there will not be much diversity in the first set of names.
Our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deter us, this should encourage us to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not stay with people in privileged groups (e.g. avoid `manels' and all-white panels, for instance).
Make sure the composition of all these groups is balanced across topics, areas, and representation across regions, gender, and race.
%rita: For academic conferences consider also to select for a talk master and PhD students and not to give voice only to senior scientist or principal investigator. lo pongo acá porque sí faltan cosas prácticas.

@@

The final list of accepted talks can also be biased. 
On the one hand, because people do not feel confident to submit their work. 
To counteract bias in submission, promote your calls beyond the usual communities and keep in touch with local groups and communities of practice to encourage submissions by their members.
Some of these communities have pre-submission mechanisms to help their members prepare and improve their submissions.%sin el ejemplo no se entiende
% 1. Clear rules on reviews: double blind, double open, kind of open, etc.
% Be transparent on the evaluation criteria
% 2. Diverse program committee (above)
% 3. Unconscious bias (test)
% 4. Curate the final list against 
%see: https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
On the other hand, proposals may be biased because of even diverse selection committees can be subjected to unconscious or implicit bias %explain!.
Unconscious bias happens when we interpret information, such as names of the authors, we assume their origin or ethnicity, their institution of origin, and we attach a positive or negative value to their work according to negative or positive stereotypes. 
To counteract this...
The review process should guarantee the fair evaluation of the submitted work, and this may  will be double blind, open, or partially open. 
In some fields, it's not possible to guarantee anonymity of the presenters, and in some and external cues such as country of origin and links to repositories. 
Create clear evaluation criteria for the submitted proposals, and share them with the authors and reviewers.
Remind the reviewers that unconscious bias exists and encourage them to take unconscious 
Selection committees need to be reminded of such biases, and be aware of them when evaluating submissions and presentations.
\cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
Activities such as scientific publication and development of software are often regarded as higher value than other roles like community building or teaching. 
These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}. Research about these topics is equally undervalued compared to other areas of knowledge production. 
Counteracting bias also means changing the standard program of the conference, proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials, // and giving visibility to the whole range of activities and practitioners that contribute to the overall topics of the conference.

%Defy the stereotypical criteria for merit and how it is traditionally measured by acknowledging these community practices and the people behind them.
You can also reframe the awards ceremony to acknowledge work beyond the traditional definition of merit/quality and include those who contributed to community building, presented teaching materials, and assumed other care-taking roles.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
%In general, people should have the option to locate themselves in a category they consider affordable, even with the possibility of a `pay what you can' approach. 
Resources permitting, allow the possibility of a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than non-free events \cite{eventbrite_ultimate_2017}. 
%Other costs can be prohibitive-> scholarships
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (the fact that some people refrain from applying because they think they don't stand a chance). 
Conferences usually ask for cover letters or applications to assign these scholarships, and this can be a time-consuming, emotionally-demanding task. 
Simplify as much as you can the process of asking for financial support; 
people who do not have the time (e.g., because they have family responsibilities) might decide not to apply if the process is overtly complicated. 
Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups. Whenever possible, facilitate ways around money transfer (e.g., book flights, hotel reservations, waive conference fees).

>>>>

% Rocío: Idea of this paragraph: non-transferability and do not restrict people
People with disabilities often say: `Nothing about us without us' [ref]; the same holds for other dimensions of diversity. The experience of people in marginalized groups facing barriers for equal participation and trying to solve them are invaluable; they cannot be replaced by good intentions from people that have not been through the same experiences \cite{costanzachockDesign2020}.
% People who have experienced exclusion have the social and technical expertise to fight against it and they will be essential in any team.%ast maybe frame this on a more positive note (ideas, problem solving)
However, do not require nor expect minoritized people to only work—or want to work—on diversity and inclusion aspects; 
everybody must have the freedom to choose in which areas of the conference they want to work. 
% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
Furthermore, do not assume that self-nomination and voting will work as mechanisms to choose leadership positions. Instead, nominate directly and offer leading positions that would normally be occupied by people from privileged groups.
Build an environment in which every person can express their opinion and pay special attention to people from systematically excluded groups.
Many people lack the institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget as well as other types of recognition (e.g. public recognition on social media, on the conference website, or during the conference).
In addition, tasks such as receiving and responding to Code of Conduct reports can be emotionally intense work and should be additionally rewarded.

@@

When choosing or inviting people for visible and valued roles in the conference––such as keynote speakers, program/scientific committee members, session chairs––, it is likely that there will not be much diversity in the first set of names.
Our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deter us, this should encourage us to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not stay with people in privileged groups (e.g. avoid `manels' and all-white panels, for instance).
Make sure the composition of all these groups is balanced across topics, areas, and representation across regions, gender, race, career stage, among other aspects.
%rita: For academic conferences consider also to select for a talk master and PhD students and not to give voice only to senior scientist or principal investigator. lo pongo acá porque sí faltan cosas prácticas. % RJ: I added career stage to the sentence above. That should be enough.

@@

The final list of accepted talks can also be biased in both submission and acceptance processes. 
People who have never participated in conferences, and especially those regarded as big and prestigious, may not feel confident to submit their work.
To counteract this self-selection bias, promote the submission call beyond the communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
Some of these communities may have pre-submission mechanisms to help their members prepare and improve their submissions (e.g. the Latin-R community has a Slack channel with the purpose of helping community members improve their abstracts for conferences, and the R-Ladies community did the same for useR! 2021).%sin el ejemplo no se entiende
% 1. Clear rules on reviews: double blind, double open, kind of open, etc.
% Be transparent on the evaluation criteria
% 2. Diverse program committee (above)
% 3. Unconscious bias (test)
% 4. Curate the final list against 
%see: https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf
% Unconscious (implicit) bias is the automatic impact of attitudes or stereotypes on our on our view of the world, actions, and decisions without our awareness (swartz 2019)
% in this study, we defined implicit bias as associations, attitudes, or beliefs that exist and enact their influence outside of an individual’s conscious awareness.
On the other hand, the review process may be biased because selection committees—even diverse ones—can be subjected to unconscious or implicit bias.
Unconscious bias happens when we interpret information, such as names of the authors, we assume their origin or ethnicity, their institution of origin, and we attach a positive or negative value to their work according to negative or positive stereotypes. Nobody is free from implicit bias [cite Ross2019].
Selection committees need to be reminded of such biases (the implicit association test can help identifying them [cite Greenwald - Measuring individual differences in implicit cognition: The implicit association test]), and
decide on the most appropriate strategy to counteract them, either with double blind, double open, or single blind reviews [cite discover cookbook].
Double blind reviews can help counteract implicit bias only if it is possible to completely anonymize the identifying information from submission. 
Double open, though more transparent, may still be subjected to implicit bias, and reviewers from minoritized groups might feel additional pressure because the reviews are signed. If this likely case, and in fields were it is not possible to guarantee anonymity of the authors during the selection process (e.g.  non-anonymized github repositories need to be examined), single blind reviews are the only option left.  
In any case, create clear evaluation criteria for the submitted proposals that do not put people from minoritized groups at disadvantage (e.g. based on their number of publications or talks), and share them with the authors and reviewers.
Advertising your practices to mitigate bias in proposal selection may also encourage people from diverse groups to submit their work, thus counteracting self selection bias.

% \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
Implicit bias also pertain the value that we give to certain activities; e.g. scientific publication and development of software are often regarded as higher value than other roles like community building or teaching. 
These roles may be equally if not more challenging and are usually assumed by women, people of color, people with disabilities, and other minoritized groups \cite{cheng2020x+, burfordHomelinessMeantHaving2020}. Research about these topics is undervalued relative to other areas of knowledge production. 
Counteracting bias also means innovating the program of the conference, proposing new thematic sessions, broadening the scope of talks, keynotes, and tutorials, and giving visibility to the whole range of activities and practitioners that contribute to the overall topics of the conference.

%Defy the stereotypical criteria for merit and how it is traditionally measured by acknowledging these community practices and the people behind them.
% You can also reframe the awards ceremony to acknowledge work beyond the traditional definition of merit/quality and include those who contributed to community building, presented teaching materials, and assumed other care-taking roles.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status).

Resources permitting, allow the possibility of a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than non-free events \cite{eventbrite_ultimate_2017}. 
%Other costs can be prohibitive-> scholarships
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (the fact that some people refrain from applying because they think they don't stand a chance). 
Conferences usually ask for cover letters or applications to assign these scholarships, and this can be a time-consuming, emotionally-demanding task. 
Simplify as much as you can the process of asking for financial support; 
people who do not have the time (e.g., because they have family responsibilities) might decide not to apply if the process is overtly complicated. 
Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups. Whenever possible, facilitate ways around money transfer (e.g., book flights, hotel reservations, waive conference fees).
